---
title: "Do users find AI-generated emails more professional than human-written ones?"
author:
    - "Bogdan Bancescu"
    - "Célia Pallier"
    - "Charles Watson Ndethi Kibaki"
    - "Chingiz Saidov"
    - "Christabel Camilleri"
date: "February 24, 2025"
format:
    pdf:
        number-sections: true
        fig-width: 8
        fig-height: 6
        keep-tex: true
        documentclass: article
        geometry: "margin=1in"
        header-includes:
          - \usepackage{microtype}
          - \sloppy
          - \setlength{\emergencystretch}{3em}
          - |
              \usepackage{etoolbox}
              \AtBeginEnvironment{quote}{\small\ttfamily}
bibliography: a2-references.bib
abstract: |
    This study investigates whether users perceive AI-generated emails as more professional than human-written ones. Using a within-subjects experimental design, participants evaluated pairs of emails on the same topic, one AI-generated and one human-written. The results indicate a slight preference for AI-generated emails in terms of professionalism, particularly among senior professionals. However, the ability to distinguish between AI and human-written emails remains imperfect, suggesting that AI has achieved a significant level of sophistication in professional communication. The study also explores the impact of professional experience and language background on these perceptions.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
```

# Introduction

In the last few years, the use of artificial intelligence (AI) in communication has increased dramatically [@brown2023ai]. Tools have been developed to generate text that mimics human writing with increasing accuracy. One task that has been greatly improved with AI is email writing. While AI tools can help gain time, reduce stress, and adopt appropriate tones, the question remains whether recipients perceive AI-generated emails as equally or more professional than those written by humans [@smith2024communication].

## Topic and Hypothesis

The research question directing this study is: "Do users find AI-generated emails more professional than human-written ones?" This question aims to understand how an email's source influences how recipients rate its professionalism.

Our hypothesis states that "AI-generated emails are more professional than human-written ones." This hypothesis suggests that the specific structure, wording, and tone used by AI in emails result in higher professionalism ratings compared to emails written by humans [@jones2024workplace].

# Methodology

## Variables

To test whether AI-generated emails are perceived as more professional than human-written ones, we defined the following variables:

- **Independent Variable (IV)**: The email source (AI-generated vs. human-written)
- **Dependent Variable (DV)**: Professionalism rating on a scale from 1 to 5
- **Control Variables**: Email topic/context (kept consistent for each pair of emails)

## Experimental Design

The experiment uses a within-subjects design where participants evaluate multiple pairs of emails, with each pair containing one AI-generated and one human-written email on the same topic [@wilson2023experimental]. This design allows for direct comparison while controlling for individual differences in rating tendencies.

```{r fig-ratings, fig.cap="Average Professionalism Ratings by Source"}
ratings_data <- data.frame(
  Source = c("AI-Generated", "Human-Written"),
  Rating = c(4.1, 3.9),
  SD = c(0.8, 0.9)
)

ggplot(ratings_data, aes(x = Source, y = Rating)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.6) +
  geom_errorbar(aes(ymin = Rating - SD, ymax = Rating + SD), width = 0.2) +
  theme_minimal() +
  labs(y = "Average Professionalism Rating", x = "Email Source") +
  ylim(0, 5)
```
## Ethical Guidelines

**Informed Consent:**  
Before participating, all individuals will be provided with clear information regarding the study's purpose and objectives. A consent form will appear at the beginning of the questionnaire, featuring an explicit consent box. This form will explain that participation is entirely voluntary, that participants can withdraw at any time without consequence, and that they are free to skip any questions they feel uncomfortable answering.

**Anonymity and Confidentiality:**  
No personal information—such as names, email addresses, ID card numbers, home addresses, gender, or religious affiliation—will be collected. All responses will remain anonymous and will be used exclusively for the purposes of this research. Data will be securely stored on a password-protected laptop, with access strictly limited to the researcher, ensuring complete confidentiality.

**Data Protection and GDPR Compliance:**  
This study adheres to the General Data Protection Regulation (GDPR) [@hoofnagle2019european], which has been in effect since 25 May 2018, with the primary goal of preserving personal data privacy. Data will be retained only as long as necessary to achieve the study's objectives and will be securely discarded once the experiment is completed. Compliance with GDPR guidelines minimizes the risk of data breaches [@voigt2017eu]. In cases of severe non-compliance, fines may reach up to 20 million euros or 2% of the previous year's global turnover, depending on which is higher. Lesser violations will be addressed on a case-by-case basis.

**Potential Risks:**  
The study poses minimal risk as it primarily involves reading emails and providing opinions. By not collecting sensitive personal information, the risk of moral or psychological harm is substantially reduced.

**Voluntary Participation and Data Sharing:**  
Participation is completely optional. Participants are free to refuse to answer any questions or withdraw from the study at any point without providing an explanation. Furthermore, no data will be exchanged with third parties, ensuring that all information remains confidential and secure throughout the research process.

## Statistical Analysis

```{r fig-experience, fig.cap="Professionalism Ratings by Professional Experience"}
experience_data <- data.frame(
  Experience = c("Students", "Early Career", "Mid-Career", "Senior"),
  AI = c(4.0, 3.3, 3.8, 4.4),
  Human = c(3.7, 4.5, 3.5, 4.2)
) %>%
  tidyr::pivot_longer(cols = c(AI, Human), 
                      names_to = "Source", 
                      values_to = "Rating")

ggplot(experience_data, aes(x = Experience, y = Rating, fill = Source)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(y = "Average Rating", x = "Professional Experience") +
  scale_fill_brewer(palette = "Set2")
```

```{r fig-accuracy, fig.cap="Source Identification Accuracy"}
accuracy_data <- data.frame(
  Category = c("Correct", "Incorrect", "Unsure"),
  Percentage = c(62.5, 18.75, 18.75)
)

ggplot(accuracy_data, aes(x = Category, y = Percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(y = "Percentage", x = "Identification Result") +
  ylim(0, 100)
```

# Results

## Statistical Findings

Our analysis revealed several key findings:

1. Overall Professionalism Ratings:
   - AI-Generated: Mean = 4.1 (SD = 0.8)
   - Human-Written: Mean = 3.9 (SD = 0.9)

2. Source Identification Accuracy:
   - Correct identification: 62.5%
   - Incorrect identification: 18.75%
   - Uncertain: 18.75%

3. Professional Background Impact:
   - Students (n=3): AI preference (4.0 vs 3.7)
   - Early Career (n=1): Human preference (4.5 vs 3.3)
   - Mid-Career (n=1): AI preference (3.8 vs 3.5)
   - Senior (n=3): AI preference (4.4 vs 4.2)

## Language and Usage Analysis

Analysis of participant demographics revealed:

1. Language Background Effect:
   - Native speakers (n=1) showed higher variation in ratings (SD = 1.1)
   - Non-native speakers (n=7) gave more consistent ratings (SD = 0.7)

2. Email Usage Impact:
   - Frequent users (multiple times/day, n=6) showed 71% source identification accuracy
   - Less frequent users (once/day, n=2) showed lower accuracy rates

# Discussion

The visualizations and statistical analysis reveal several interesting patterns:

1. AI-generated emails consistently received slightly higher professionalism ratings, supporting our initial hypothesis [@davis2024ai].
2. Professional experience influences rating patterns, with senior professionals showing the strongest preference for AI-generated content.
3. Source identification accuracy varies significantly, suggesting that AI writing has become sophisticated enough to challenge human perception [@miller2024natural].
4. Rating consistency is higher for AI-generated emails, possibly due to their more standardized structure and tone [@taylor2024consistency].

## Limitations

Several limitations should be considered when interpreting these results:

1. Sample Size: The relatively small sample size (n=8) limits the generalizability of findings.
2. Professional Distribution: Uneven distribution across professional categories affects comparative analysis.
3. Email Topics: Limited to four scenarios, which may not represent all professional communication contexts.

## Future Research Directions

Future studies could address these limitations by:

1. Expanding the sample size and ensuring even distribution across professional categories.
2. Including a wider variety of email contexts and communication scenarios.
3. Conducting longitudinal studies to track changes in perception over time [@chen2024longitudinal].

# Conclusion

Our findings suggest a slight preference for AI-generated emails in terms of perceived professionalism, particularly among senior professionals. However, the ability to distinguish between AI and human-written emails remains imperfect, indicating that AI has achieved a significant level of sophistication in professional communication.

# Appendices

## Appendix A: Email Samples

### Service Update Announcement (Epic Telecom)

#### AI-Generated Version
> Dear Valued Customer,
> 
> We are excited to announce a series of service updates from Epic designed to provide you with faster and more reliable connectivity. Please review the key highlights below:
> 
> * Next-Generation 5G Network: Experience lightning-fast speeds and reduced latency as Epic rolls out its enhanced 5G coverage across urban and suburban areas.
> * Upgraded Broadband Packages: Enjoy our newly designed broadband packages offering higher speeds and flexible data options tailored to your household needs.
> * Exclusive Consumer Offers: Take advantage of limited-time promotions, including discounted rates on bundled mobile and home internet services.
> * Enhanced Customer Support: Benefit from our upgraded support system that ensures faster response times and personalized service.
> 
> For further details on these exciting updates and special offers, please visit our website or contact our support team.
> 
> Thank you for choosing Epic as your trusted telecom provider.
> 
> Sincerely,
> The Epic Team

#### Human-Written Version
> Hi there,
> 
> We've got some exciting news from Epic just for you! Our plans are getting a major upgrade, and we want you to be the first to know:
> 
> * Our new 5G network is rolling out soon—get ready for super-fast speeds and a smoother connection.
> * We've revamped our broadband plans to give you more speed and better deals. Enjoy more streaming, gaming, and browsing without interruptions.
> * And because we appreciate you, Epic is offering some sweet discounts on bundled plans for a limited time.
> 
> If you're curious to learn more or just want to chat about which plan suits your needs best, drop us a reply or check out our website.
> 
> Thanks for being a part of the Epic community, and we can't wait for you to experience these new benefits!
> 
> Cheers,
> The Epic Team

### Order Status Notification

#### AI-Generated Version
> Subject: Follow-Up on Order No. 64235993
> 
> Dear Ms. Durand,
> 
> I hope this message finds you well. Regarding your order No. 64235993, I wanted to inform you that product 56463 was discontinued last year and is no longer available.
> 
> However, our company does offer an alternative product. While it is not an exact match, it serves a similar purpose. Please let us know if you would like us to include the alternative in your order or if you prefer to remove this item entirely.
> 
> We look forward to your decision so we can proceed accordingly.
> 
> Best regards,
> John Doe

#### Human-Written Version
> Subject: Issue in the order n°64235993: Discontinued Product 56463
> 
> Dear Ms Durand,
> 
> I reviewed your order request but it seems there was a mistake with the product 56463. Indeed, this product was discontinued a year ago, so unfortunately, we won't be able to include it in your order. We have alternative products, but they are different in some aspects.
> Before we proceed with your order, could you please inform us if you would like an alternative or if you prefer to remove the product from the order?
> 
> Best regards,
> John Doe

### Job Application Response

#### AI-Generated Version
> Dear Mr. Constantin,
> 
> Thank you for taking the time to apply for the Backend Developer position at Major and for sharing your skills and experience with us. We truly appreciate your interest in joining our team.
> 
> After careful consideration, we have decided to move forward with another candidate for this role. This decision was not easy, as we received applications from many qualified professionals, including yourself.
> 
> We want to recognize the effort you put into the application process and encourage you to stay connected with us for future opportunities that match your expertise. Please feel free to apply again if a suitable role arises.
> 
> Wishing you all the best in your job search and future career endeavors.
> 
> Best regards,
> Anna HR Generalist

#### Human-Written Version
> Dear Mr. Constantin,
> 
> Thank you for your interest in the Backend Developer position at Major. We appreciate the time and effort you invested in submitting your application and CV.
> 
> After careful consideration, we regret to inform you that we have decided to move forward with other candidates whose qualifications more closely align with our current needs and requirements.
> 
> We encourage you to apply for future openings that match your skills and experiences, as we were impressed by your qualifications and believe you could be a valuable asset to our team.
> 
> Once again, thank you for your interest in Major. We wish you all the best in your job search and future professional endeavors.
> 
> Best regards,
> Anna HR Generalist

### Network Infrastructure Update

#### AI-Generated Version
> Dear Colleagues,
> 
> I'm writing to share an important update about changes to our WIFI network infrastructure. As part of our ongoing efforts to enhance network security and optimize bandwidth usage, we will be transitioning to a dual-network system starting September 7, 2024.
> 
> We've created two dedicated networks to better serve different device types:
> 
> AquilaNet will continue to be our primary network for all company-issued devices, ensuring secure access to our corporate resources.
> 
> AquilaGuest will be a separate network specifically designed for personal devices and visitors. This includes mobile phones, tablets, smartwatches, and any non-company devices.
> 
> On September 7, all personal devices currently connected to AquilaNet will need to switch to the AquilaGuest network. To make this transition smooth, simply connect to:
> 
> Network: AquilaGuest  
> Password: Guest1234!
> 
> This network segmentation will help us better manage our bandwidth resources while maintaining robust security standards. The IT support team is here to help should you encounter any connectivity issues during or after the transition.
> 
> We appreciate your cooperation in implementing these important security measures.
> 
> Best regards,
> Lucius Verus  
> Network Infrastructure Team

#### Human-Written Version
> Hello all,
> 
> In an effort to further strengthen our network security posture and to more efficiently utilize our internet bandwidth, we have segmented our WIFI network into two networks: AquilaNet and AquilaGuest.
> 
> Our company issued work devices will continue to reside on AquilaNet.
> 
> However effective 09/07/2024, our personal devices (mobile devices, smart watches, tablets, etc) and office guests' devices will be disconnected from AquilaNet and will only be able to connect to AquilaGuest.
> 
> To remain connected on your personal device, connect to the AquilaGuest using this password: Guest1234!
> 
> Feel free to reach out to me, in case you encounter any issues connecting to either network.
> 
> Happy connecting.
> 
> Sincerely,
> Lucius Verus

## Appendix B: Survey Instrument
The survey was conducted using Google Forms and is available at the following link:
[Email Professionalism Study Questionnaire](https://docs.google.com/forms/d/e/1FAIpQLSeb-7fgi9-pYoLTwQS9BhUcwpGYxNoAVz20g9WyYCUdVvJXDQ/viewform?usp=header)

The survey included:

1. Demographic Questions:
   - Age group
   - Professional background
   - Industry/sector
   - Email usage frequency
   - Primary language

2. For each email:
   - Source identification (AI/Human/Unsure)
   - Professionalism rating (1-5 scale)

## Appendix C: Statistical Methods

Detailed statistical procedures included:

1. Descriptive Statistics:
   - Mean and standard deviation calculations
   - Frequency distributions
   - Cross-tabulations

2. Inferential Statistics:
   - Chi-square tests for independence
   - Paired t-tests for rating comparisons
   - Correlation analyses

# References {.unnumbered}

::: {#refs}
:::